---
title: "Index"
author: "Chris Drew"
format: html
---

#Part 1 Questions
```{r}
# Load necessary libraries
library(tidyverse)
library(ggpubr)
library(lterdatasampler)
library(car)

# Clean the data
sally_clean <- and_vertebrates |> 
  filter(species == "Coastal giant salamander") |> 
  drop_na(unittype, section)

# contingency table
cont_table <- table(sally_clean$section, sally_clean$unittype)

# Print contingency table
cont_table

chisq_test <- chisq.test(cont_table)

chisq_test$expected

# chi-square test
chisq_test_result <- chisq.test(cont_table)

chisq_test_result

# Visualize
sally_clean |> 
  count(unittype, section) |>  # Calculate counts of unittype by section
  ggbarplot(x = 'unittype', y = 'n', fill = 'section', 
            palette = c("#00AFBB", "#E7B800"), add = "mean_se") +
  labs(title = "Unit Type by Forest Type for Coastal Giant Salamander", 
       y = "Count", x = "Unit Type")

# Clean 
sally_clean <- and_vertebrates |> 
  filter(species == "Coastal giant salamander") |> 
  drop_na(weight_g)

# Separate the data 
cc_weight <- sally_clean |> 
  filter(section == "CC") |> 
  pull(weight_g)

og_weight <- sally_clean |> 
  filter(section == "OG") |> 
  pull(weight_g)

shapiro.test(cc_weight)  # Normality for Clear Cut
shapiro.test(og_weight)  # Normality for Old Growth

leveneTest(weight_g ~ section, data = sally_clean)

t_test_result <- t.test(cc_weight, og_weight, var.equal = TRUE)

# Report the t-test result
t_test_result

# Visualize the weight data for both Clear Cut and Old Growth sections
ggarrange(
  gghistogram(cc_weight, title = "Clear Cut Forest - Coastal Giant Salamander Weight"),
  gghistogram(og_weight, title = "Old Growth Forest - Coastal Giant Salamander Weight")
)

# Filter for Cutthroat trout and remove missing values for length_2_mm and weight_g
trout_clean <- and_vertebrates %>%
  filter(species == "Cutthroat trout") %>%
  drop_na(length_2_mm, weight_g)

shapiro.test(trout_clean$length_2_mm)
shapiro.test(trout_clean$weight_g)


cor.test(trout_clean$length_2_mm, trout_clean$weight_g, method = "pearson")

cor.test(trout_clean$length_2_mm, trout_clean$weight_g, method = "spearman")

ggscatter(trout_clean, x = "length_2_mm", y = "weight_g", 
          color = "species", add = "reg.line", 
          conf.int = TRUE, palette = "jco") +
  labs(title = "Correlation between Body Length and Weight for Cutthroat Trout", 
       x = "Body Length (mm)", y = "Body Weight (g)")

```
#Part 2 Questions
```{r}
library(tidyverse)
library(ggpubr)
library(broom)
library(car)
library(flextable)
```


```{r}
data("pie_crab")
glimpse(pie_crab)
summary(pie_crab)

# Sample size per site
count(pie_crab, site)

# Visualizing crab size distribution per site
ggplot(pie_crab, aes(x = site, y = size, col = site)) +
  geom_boxplot() +
  geom_jitter(size = 0.25) +
  theme(legend.position = "none")

# Check normality using Shapiro-Wilk test
norms <- pie_crab |> 
  nest(data = -site) |>
  mutate(Shapiro = map(data, ~ shapiro.test(.x$size)),
         n = map_dbl(data, nrow),
         glance_shapiro = map(Shapiro, broom::glance)) |>
  unnest(glance_shapiro)

flextable::flextable(dplyr::select(norms, site, n, statistic, p.value)) |>
  flextable::set_caption("Shapiro-Wilk normality test for size at each site")

# Run ANOVA
res_aov <- aov(size ~ site, data = pie_crab)
summary(res_aov)

# Check residuals' normality
gghistogram(res_aov$residuals)
shapiro.test(res_aov$residuals)

# Test for equal variances
leveneTest(size ~ site, data = pie_crab)

# Perform Welch's ANOVA
oneway.test(size ~ site, data = pie_crab, var.equal = FALSE)

# Subset data for selected sites
pie_sites <- pie_crab |> 
  filter(site %in% c("GTM", "DB", "PIE"))

# Check for equal variance in subset
leveneTest(size ~ site, data = pie_sites)

# Run ANOVA for subset
pie_anova <- aov(size ~ site, data = pie_sites)
summary(pie_anova)

# Tukey's HSD post-hoc test
TukeyHSD(pie_anova)

# Linear regression: Testing Bergmann’s rule
pie_lm <- lm(size ~ latitude, data = pie_crab)

# View regression results
summary(pie_lm)



```

```{r}
str(pie_crab)
library(ggplot2)
library(dplyr)

# Ensure site is ordered by latitude
pie_crab <- pie_crab %>%
  mutate(site = reorder(site, latitude))

# Create the boxplot
boxplot_figure <- ggplot(pie_crab, aes(x = site, y = size)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(
    x = "Site (Ordered by Latitude)",
    y = "Carapace Width (mm)",
    title = "Carapace Width by Site"
  ) +
  theme_minimal()

# Save the plot
ggsave("img/crab_boxplot.png", boxplot_figure, width = 6, height = 4, dpi = 300)

# Display the plot
print(boxplot_figure)



```
#Report:Latitude was a significant predictor of carapace width (p < 0.001), while air temperature standard deviation (p = 0.173) and water temperature standard deviation (p = 0.325) were not significant.

```{r}
# Load necessary packages
library(tidyverse)
str(pie_crab)
model <- lm(size ~ water_temp_sd, data = pie_crab)
summary(model)
library(ggplot2)

ggplot(pie_crab, aes(x = water_temp_sd, y = size)) +
  geom_point() +
  geom_smooth(method = "lm", col = "red") +
  labs(x = "Annual Variation in Water Temperature (SD)", y = "Carapace Width (Size)") +
  theme_minimal()
ggsave("carapace_width_vs_water_temp_sd.png")


```
#Report:The plot shows the relationship between water temperature standard deviation (water_temp_sd) and carapace width, with data points scattered around the line of best fit. The regression reveals no large effect of water_temp_sd on carapace width, the p-value for the water_temp_sd coefficient is 0.528, showing no significant relationship.
```{r}
# Select the predictor variables
predictors <- pie_crab[, c("latitude", "air_temp_sd", "water_temp_sd")]

cor_matrix <- cor(predictors)

# Display the correlation matrix
cor_matrix
# Fit the multiple linear regression model
model_multiple <- lm(size ~ latitude + air_temp_sd + water_temp_sd, data = pie_crab)

# Summarize
summary(model_multiple)



```
#Report: Latitude significantly influences carapace width (p < 2e-16), while air temperature variability (p = 0.1730) and water temperature variability (p = 0.3254) do not significantly affect carapace width.
